---
title: "Machine Learing in Predicting Bitcoin Price"
author: "Yunting Chiu"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Install the Libraries
```{r, message=FALSE}
library(tidyverse)
library(tidymodels)
```

#REPORT ABSTRACT:
As Wall Street giants, retail investors, and aspiring cryptocurrency trailblazers continue to flood the cryptocurrency market, the ability to predict the volatility of cryptocurrency stocks has proven to be increasingly invaluable. In this report, we detail our methodology that applies statistical machine learning techniques to predict the direction of Bitcoin stocks. Our work also aims to build upon previous research conducted in anticipating trends within cryptocurrency using statistical machine learning methods. To predict whether the direction of Bitcoin stocks will increase or decrease on a given date, we employ Linear Discriminant Analysis (LDA), Quadratic Discriminant Analysis (QDA), K-Nearest Neighbors (KNN), Logistic Regression Analysis, Random Forest, and Decision Trees. We also perform statistical analyses of our dataset using simple linear regression, multiple linear regression, and summary statistics, and visualize these metrics to gain a more comprehensive understanding of the variables that may contribute to deciding the direction of Bitcoin stocks on a given day. Finally, we analyze our results and discuss opportunities for future work and research. 

# TextBook Example
```{r}
library(ISLR)
Smarket
```

# BitCoin Data Set Pre-Processing
- https://financeformulas.net/Total-Stock-Return.html
- https://dplyr.tidyverse.org/reference/lead-lag.html
```{r}
BTC <- read_csv("./data/BTC_USD_2013-10-01_2021-06-13-CoinDesk.csv")
BTC %>%
  rename(Closing_Price = "Closing Price (USD)",
         Open_Price = "24h Open (USD)") %>%
  mutate(Lag1 = (lag(Closing_Price, n = 1) - lag(Open_Price, n = 1)) / lag(Open_Price, n = 1)*100,
         Lag2 = (lag(Closing_Price, n = 2) - lag(Open_Price, n = 2)) / lag(Open_Price, n = 2)*100,
         Lag3 = (lag(Closing_Price, n = 3) - lag(Open_Price, n = 3)) / lag(Open_Price, n = 3)*100,
         Lag4 = (lag(Closing_Price, n = 4) - lag(Open_Price, n = 4)) / lag(Open_Price, n = 4)*100,
         Lag5 = (lag(Closing_Price, n = 5) - lag(Open_Price, n = 5)) / lag(Open_Price, n = 5)*100,
         Return_Today = (Closing_Price - Open_Price) / Open_Price*100,
         Direction_Today = ifelse(Return_Today > 0, "Up", "Down"),
         Direction_Today = as.factor(Direction_Today)
         ) -> BTC_tidied 

names(BTC_tidied)
head(BTC_tidied)
tail(BTC_tidied)
```

```{r}
View(BTC_tidied)
```

```{r}
BTC_tidied %>%
  drop_na() -> BTC_tidied
```

#Statistical Analysis
```{r}
summary(BTC_tidied)
```

Multiple Linear Regression Model 1: 
```{r}
btc_mlc <- lm(Return_Today ~ Lag1 + Lag2 + Lag3 + Lag4 + Lag5, data = BTC_tidied)
btc_mlc
```

```{r}
summary(btc_mlc)
```

```{r}
plot(btc_mlc)
```

Multiple Regression Model 2 (with added pre-processing for this model)
```{r}
btc_mlc2 <- lm(as.character(Direction_Today ~ Lag1 + Lag2 + Lag3 + Lag4 + Lag5, data = BTC_tidied))
btc_mlc2 
#Note: Maybe I'm not implementing this code correctly, but it seems R won't register "Direction_Today" as a variable here. I'd be okay with us omitting this second multiple regression model altogether and only keeping the first one that worked above, so let me know what you think! I also think using the "Return_Today" as the response variable for the rest of the simple linear regression models should be okay, but let me know if you agree. If you agree with me on both points, I think this code should be set to merge with yours! 
```

```{r}
btc_mlc2 <- lm(Direction_Today ~ Lag1 + Lag2 + Lag3 + Lag4 + Lag5, data = BTC_tidied)
btc_mlc2
```

```{r}
summary(btc_mlc2)
```

```{r}
plot(btc_mlc2)
```

Simple Linear Regression Model 1:
```{r}
btc_slr <- lm(Return_Today ~ Lag1, data = BTC_tidied)
btc_slr
```

```{r}
summary(btc_slr)
```

Simple Linear Regression Model 2:
```{r}
btc_slr2 <- lm(Return_Today ~ Lag2, data = BTC_tidied)
btc_slr2
```

```{r}
summary(btc_slr2)
```

Simple Linear Regression Model 3:
```{r}
btc_slr3 <- lm(Return_Today ~ Lag3, data = BTC_tidied)
btc_slr3
```

```{r}
summary(btc_slr3)
```

Simple Linear Regression Model 4:
```{r}
btc_slr4 <- lm(Return_Today ~ Lag4, data = BTC_tidied)
btc_slr4
```

```{r}
summary(btc_slr4)
```

Simple Linear Regression Model 5:
```{r}
btc_slr5 <- lm(Return_Today ~ Lag5, data = BTC_tidied)
btc_slr5
```

```{r}
summary(btc_slr5)
```

#Decision Tree and Random Forest
Decision Tree Model 1: 
```{r}
bc_split <- initial_split(BTC_tidied)
set.seed(1234)
bc_train <- training(bc_split)
bc_test <- testing(bc_split)
```

```{r}
decision_tree_rpart_spec <- 
  decision_tree() %>%
  set_engine('rpart') %>%
  set_mode('classification')
```

```{r}
dt_fit <- fit(decision_tree_rpart_spec, Direction_Today ~., data = bc_train)
dt_fit
```

```{r}
library(rpart.plot)
rpart.plot(dt_fit$fit)
```

Decision Tree Model 2: Hyperparameter
```{r}
decision_tree_rpart_spec_lambda <-
  decision_tree(cost_complexity = 0.5) %>%
  set_engine('rpart') %>%
  set_mode('classification')
  
dt_fit_lambda2 <- fit(decision_tree_rpart_spec_lambda, Direction_Today ~., data = bc_train)
dt_fit_lambda2
```

```{r}
rpart.plot(dt_fit_lambda2$fit)
```

Decision Tree Model 3: 
```{r}
decision_tree_rpart_spec_lambda <-
  decision_tree(cost_complexity = 3) %>%
  set_engine('rpart') %>%
  set_mode('classification')
  
dt_fit_lambda3 <- fit(decision_tree_rpart_spec_lambda, Direction_Today ~., data = bc_train)
dt_fit_lambda3
```

```{r}
rpart.plot(dt_fit_lambda3$fit)
```

Decision Tree Model 4: 
```{r}
decision_tree_rpart_spec_lambda <-
  decision_tree(cost_complexity = 1.5) %>%
  set_engine('rpart') %>%
  set_mode('classification')
  
dt_fit_lambda4 <- fit(decision_tree_rpart_spec_lambda, Direction_Today ~., data = bc_train)
dt_fit_lambda4
```

```{r}
rpart.plot(dt_fit_lambda4$fit)
```

```{r}
library(vip)
vi(dt_fit)
vip(dt_fit)
```

Random Forest
```{r}
rand_forest_ranger_spec <- 
  rand_forest() %>%
  set_engine('ranger', importance = "impurity") %>%
  set_mode('classification')
rand_forest_ranger_spec
```

```{r}
library(ranger)
set.seed(4321)
rf_fit <- fit(rand_forest_ranger_spec, Direction_Today ~., data = bc_train)
rf_fit
```

```{r}
vi(rf_fit$fit)
vip(rf_fit$fit)
```

#Data Visualization 
```{r, message = FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

Model 1: Scatter Plot of Date vs Return Today
```{r}
ggplot(data=BTC_tidied) +
  geom_point(mapping = aes(x = Date, y= Return_Today), color = "purple")
```

Model 2: Bar Graph of Date vs Return Today with Return_Today Fill
```{r}
ggplot(data = BTC_tidied) +
  geom_bar(mapping = aes(x = Date, y = Return_Today, fill = Return_Today), stat = "identity")
```

Model 3: Variation of Bar Graph of Date vs Return Today with Date Fill
```{r}
ggplot(data = BTC_tidied) +
  geom_bar(mapping = aes(x = Date, y = Return_Today, fill = Date), stat = "identity")
```

Model 4: Scatter Plot of Date vs Opening Price
```{r}
ggplot(data=BTC_tidied) +
  geom_point(mapping = aes(x = Date, y= Open_Price), color = "purple") 
```

Model 5: Scatter Plot of Date vs Closing Price
```{r}
ggplot(data=BTC_tidied) +
  geom_point(mapping = aes(x = Date, y= Closing_Price), color = "purple") 
```

Model 6: Pie Chart - Frequency of Ups vs Downs
```{r}
cd<- ggplot(BTC_tidied, aes(x="", y = Return_Today, fill=Direction_Today))+
  geom_bar(width = .5, stat = "identity") 
cd

pie <- cd + coord_polar("y", start=0) 
pie
```


