---
title: "STAT-627 Assignment 1"
author: "Yunting Chiu"
date: "`r Sys.Date()`"
output:
  html_document: 
    number_sections: yes
    highlight: pygments
    theme: cerulean
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment 1
Each exercise is worth 16 points.

## Exercise 1
For each of parts (a) through (d), indicate whether we would generally expect the performance of a flexible statistical learning method to be better or worse than an inflexible method. Justify your answer.

Before we answering the questions, we should know a inflexible method is a simple method; a flexible method is a complex method.

The sample size n is extremely large, and the number of predictors p is small.

- **Better**. The large number of observations is better because we can know more about the detailed correlation between them. Also, the number of small predictors and large sample sizes may help to avoid a overfitting problem and reduce the bias.

The number of predictors p is extremely large, and the number of observations n is small.

- **Worse**. It is possible that not all predictors have a significant effect on the response variable. Overfitting and spurious correlation are phenomenons that affect data models with a large number of predictors, in which the data model performs well on training data but badly on test data. Furthermore, small sample sizes and too many predictors will result in high variance.

The relationship between the predictors and response is highly non-linear.

- **Better**. If there are too many limits, just a few types of trends are conceivable, which may not convey the underlying nonlinear relationship. A better fit will result from having more degrees of freedom. Therefore, non-linearity allows predictors to be better fitted with the dependent variable.

The variance of the error terms, is extremely high.

- **Worse**. Because of the high variance, an algorithm may model the random noise training data rather than the expected outputs. High error terms would decrease the model's performance.

### References:
1. https://www.kdnuggets.com/2017/04/must-know-fewer-predictors-machine-learning-models.html
2. https://stats.stackexchange.com/questions/69237/flexible-and-inflexible-models-in-machine-learning
3. https://becominghuman.ai/machine-learning-bias-vs-variance-641f924e6c57

## Exercise 2
Describe the difference between a parametric and non-parametric statistical learning approach. What are the advantages of a parametric approach to regression or classification (as opposed to a non-parametric approach)? What are its disadvantages?



## Exercise 3
Carefully explain the the difference between the KNN classifier and KNN regression methods. Name a downside when using this model on very large data.



## Exercise 4
Suppose we have a data set with five predictors, $X1$= GPA, $X2$= extracurricular activities (EA), $X3$= Gender (1 for Female and 0 for Male), $X4$= Interaction between GPA and EA, and $X5$= Interaction between GPA and Gender. The response is starting salary after graduation (in thousands of dollars). Suppose we use least squares to fit the model, and get $\beta_0$ = 50, $\beta_1$ = 20, $\beta_2$ = 0.07, $\beta_3$ = 35, $\beta_4$ = 0.01, $\beta_5$ = −10.

The regression equation should be:
$$
\hat{salary} = 50 + 20GPA + 0.07EA + 35Gender + 0.01GPA*EA -10GPA*Gender
$$

As `Gender` is a dummy variable, the equation for females and males should be:

**Female**: gender is 1
$$
\hat{salary} = 50 + 20GPA + 0.07EA + 35 + 0.01GPA*EA -10GPA*Gender
$$
$$
\hat{salary} = 85 + 10GPA + 0.07EA + 0.01GPA*EA
$$

**Male**: gender is 0
$$
\hat{salary} = 50 + 20GPA + 0.07EA + 0.01GPA*EA 
$$
In other words, the difference between female and male is:
$$
female - male = (85 + 10GPA + 0.07EA + 0.01GPA*EA) - (50 + 20GPA + 0.07EA + 0.01GPA*EA)
$$
$$
female - male = 35 - 10GPA
$$
a. Which answer is correct, and why?\
    1. For a fixed value of EA and GPA, males earn more on average than females.\
    2. For a fixed value of EA and GPA, females earn more on average than males.\
    3. For a fixed value of EA and GPA, males earn more on average than females provided that the GPA is high enough.\
    4. For a fixed value of EA and GPA, females earn more on average than males provided that the GPA is high enough.

- (3) is correct. Because the expected salary of female subtract male is 35 - 10 * GPA, we can know the female earn more **35 - 10 * mean(GPA)** than male. According to the **female-male equation**, as the average GPA rises, the female will earn less. In conclusion, men will earn more than women if their GPA is higher.

b. Predict the salary of a female with EA of 110 and a GPA of 4.0.

- 137,100 dollars.
```{r}
EA <- 110
GPA <- 4
f_salary <- (85 + 10*GPA + 0.07*EA + 0.01*GPA*EA) * 1000 # in thousand dollars
f_salary 
```


c. True or false: Since the coefficient for the GPA/EA interaction term is very small, there is very
little evidence of an interaction effect. Justify your answer.

- This is **false**. A small value of coefficient does not imply that the interaction term has a minor influence. The statistical significance of the coefficient can be determined by looking at the p-value in the coefficient table.

## Exercise 5
This question should be answered using the `biomass` data set.
```{r}
library(tidyverse)
library(tidymodels)
data("biomass")
biomass %>%
  head()
```
a. Fit a multiple regression model to predict `HHV` using `carbon`, `hydrogen` and `oxygen`.
```{r}
# create a parsnip specification
linear_reg() %>%
  set_mode("regression") %>%
  set_engine("lm") -> lm_spec

# fit the model using tidymodel
lm_spec %>%
  fit(HHV ~ carbon + hydrogen + oxygen, data = biomass) -> lm_fit

lm_fit %>%
  pluck("fit") %>%
  summary()

# Another way to see the summary table
# tidy(lm_fit)

# predict HHV variable by using the `lm_fit` linear model
predict(lm_fit, new_data = biomass) 
```

b. Provide an interpretation of each coefficient in the model. Be careful, note the values `Cruise` is able to take.
```{r}
lm_fit %>%
  pluck("fit") %>%
  summary()
```

c. Write out the model in equation form.
$$
\hat{HHV} = 1.0456860 + 0.3478508carbon + 0.2430900hydrogen - 0.0003767oxygen
$$

d. For which the predictors can you reject the null hypothesis H0:$\beta_j = 0$?

Hypothesis test: H0: $\beta j=0$ vs Ha: $\beta j \neq{0}$

The p-value of $b_1$ and $b_2$ is 2e-16 and 0.0000986, so we have evidence to reject the null hypothesis in favor of the alternative hypothesis, meaning that $b_1$ and $b_2$ are not equal to 0. Thus, predictor `carbon` and `hydrogen` have an impact on `HHV`, when holding `oxygen` constant.

e. On the basis of your response to the previous question, fit a smaller model that only uses the predictors for which there is evidence of association with the outcome.

We consider removing the predictor that does not meet the significance level based on the results of 5a. That is, we should eliminate the variable `oxygen` (because of no impact) then re-fit the linear model. 
```{r}
lm_spec %>%
  fit(HHV ~ carbon + hydrogen, data = biomass) -> lm_new_fit

lm_new_fit$fit %>%
  summary()
```

f. How well do the models in (a) and (e) fit the data? How big was the effect of removing the predictor?



