---
title: "STAT-627 Assignment 2"
author: "Yunting Chiu"
date: "`r Sys.Date()`"
output:
  html_document: 
    number_sections: yes
    highlight: pygments
    theme: cerulean
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 1 (10 points)

Suppose we collect data for a group of students in a statistics class with variables $X_1$ = hours studied, $X_2$ = undergrad GPA, and $Y$ = receive an A. We fit a logistic regression and produce estimated coefficient, $\hat{\beta}_0=-6$, $\hat{\beta}_1=0.05$, $\hat{\beta}_2=1$.

The equation of logistic regression is:
$$
p(X) = \frac{e^{\beta_0 + \beta_1X_1 + … + \beta_pX_p}}{1 + e^{\beta_0 + \beta_1X_1 + … + \beta_pX_p}}
$$
a) Estimate the probability that a student who studies for 40 hours and has an undergrad GPA of $3.5$ gets an A in the class.
 
According to the question, $X_1$ is 40 hours and $X_2$ is $3.5$ so the probability will be:
$$
\hat{p}(X) = \frac{e^{-6 + 0.05X_1 + 1X_2}}{(1 + e^{-6 + 0.05X_1 + 1X_2})} = 0.3775.
$$

`e` is an Euler's number, which is close to 2.7182818284590452353602874713527.... Using `exp()` in the console, the probability of getting an A will be 0.3775
$$
\hat{p}(X) = \frac{e^{-6 + 0.05*40 + 1*3.5}}{(1 + e^{-6 + 0.05*40 + 1*3.5})} = 0.3775406
$$

b) How many hours would that student in part (a) need to study to have a 50% chance of getting an A in the class?

Now, we don't know how many studying hour is, we need to find out $X_1$. 
$$
\frac{e^{-6 + 0.05X_1 + 1*3.5}}{(1 + e^{-6 + 0.05X_1 + 1*3.5})} = 0.5
$$
```{r}

```

# Exercise 2 (10 points)

Suppose that we take a data set, divide it into equally-sized training and test sets, and then try out two different classification procedures. First, we use logistic regression and get an error rate of 20% on the training data and 30% on the test data. Next, we use 1-nearest neighbors (i.e. $K = 1$) and get an average error rate (averaged over both test and training data sets) of 18%. Based on these results, which method should we prefer to use for classification of new observations? Why?

# Exercise 3 (15 points)
You will in this exercise examine the differences between LDA and QDA.

a) If the Bayes decision boundary is linear, do we expect LDA or QDA to perform better on the training set? On the test set?

b) If the Bayes decision boundary is non-linear, do we expect LDA or QDA to perform better on the training set? On the test set?

c) In general, as the sample size n increases, do we expect the test prediction accuracy or QDA relative to LDA to improve, decline, or be unchanged? Why?

d) True or False: Even if the Bayes decision boundary for a given problem is linear, we will probably achieve a superior test error rate using QDA rather than LDA because QDA is flexible enough to model a linear decision boundary. Justify your answer.

Ans: False. 

## References
- http://uc-r.github.io/discriminant_analysis