---
title: "K-means Clustering"
author: "Yunting Chiu"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: cerulean
    highlight: zenburn
---
```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Palmer Station Penguin Data 
We will be using the [palmerpenguins](https://allisonhorst.github.io/palmerpenguins/) data set for this lab.

We will also be needing to load the [broom](https://broom.tidymodels.org/) package

```{r}
library(palmerpenguins)
library(broom)
library(tidymodels)
library(tidyverse)
# our data set
head(penguins)
```
a. Transform the data set into a matrix using two of the numeric variables
```{r}
# tidyverse style
penguins %>%
  select(bill_length_mm, bill_depth_mm) %>%
  drop_na() %>%
  as.matrix() -> pen2
pen2 %>%
  nrow() # the matrix has 342 observations
# Base R style
penn <- as.matrix(penguins[, c("bill_length_mm", "bill_depth_mm")])
```
# 3 Clusters
b. Perform k-means using 3 clusters
- The first to third observations is 85, 116, 141, respectively.
- The mean of each cluster has shown below.
- Each vector aims to partition n observations into k clusters in which each observation belongs to the cluster with the nearest mean.
```{r}
set.seed(1234)
pen2_kmeans <- kmeans(pen2, centers = 3) # perform k-means clustering on a data matrix
pen2_kmeans
```

c. Look at the result object with `summary()`, `names()`, and `str()`.
-  The `summary()` shows the elements in `pen2_kmeans`.
```{r}
summary(pen2_kmeans)
```
We also can found the attributes by using `names()`.

- `cluster`: A vector of integers (from 1:k) indicating the cluster to which each point is allocated.
- `centers`: A matrix of cluster centers
- `totss`: The total sum of squares.
- `withinss`: Vector of within-cluster sum of squares, one component per cluster.
- `tot.withinss`: Total within-cluster sum of squares, i.e.`sum(withinss)`.
- `betweenss`: The between-cluster sum of squares, i.e.`totss-tot.withinss`.
- `size`: The number of points in each cluster (observations).
- `iter`: The number of (outer) iterations.
- `ifault`: integer: indicator of a possible algorithm problem -- for experts.
```{r}
names(pen2_kmeans) # between within sum of square
```

For example, I want to see how many iterations in `pen2_kmeans`, I will carry out:
```{r}
pen2_kmeans$iter
```

`str()` displays the specific structure of an object.
```{r}
str(pen2_kmeans)
```

d. Use `augment()`, `glance()` and `tidy()` to extract information from the model
- `tidy()` generates a readable output.
```{r}
tidy(pen2_kmeans)
```
`glance()` shows the simple over view of the model.
```{r}
glance(pen2_kmeans)
```
`augment()` saves the results in the existing data set, automatically generating a new column.
```{r}
augment(pen2_kmeans, data = pen2)
```
Or we can select more variables from the original data and attach the results as a new data set.
```{r}
penguins %>%
  select(species, island, year, bill_length_mm, bill_depth_mm) %>%
  drop_na() -> new_penguins
augment(pen2_kmeans, data = new_penguins) -> new_clusters
head(new_clusters)
```

e. Plot the clusters with your package of choice

Showing 3 clusters in a scatterplot
```{r}
new_clusters %>%
  ggplot(aes(bill_length_mm, bill_depth_mm, color = .cluster)) +
  geom_point() +
  theme_bw()
```

Fix the x- and y-axes and enlarge the data points of cluster means.
```{r}
new_clusters %>%
  ggplot(aes(bill_length_mm, bill_depth_mm, color = .cluster)) +
  geom_point() +
  geom_point(aes(color = cluster), size = 10, data = tidy(pen2_kmeans)) +
  coord_fixed()
```

Add a `species` variable to legend.
```{r}
new_clusters %>% 
  ggplot(aes(bill_length_mm, bill_depth_mm, color = species)) +
  geom_point() +
  theme_dark()
```

Count species vs cluster sampling 
```{r}
new_clusters %>%
  count(.cluster, species) %>%
  ggplot(aes(.cluster, species, size = n)) +
  geom_point() +
  ggtitle("Species vs Cluster")
```
# 5 Clusters

f. Return The previous steps with more variables and different values of $K$

**Perform k-means using 5 clusters**: We now add two more variables `flipper_length_mm` and `body_mass_g` to the matrix, and save it as `pen3`.
```{r}
pen3 <- penguins %>%
  select(bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g) %>%
  drop_na() %>%
  scale() %>% # will calculate the mean and standard deviation of the entire vector
  as.matrix()
head(pen3)
```


```{r}
set.seed(1)
pen3_kmeans <- kmeans(pen3, centers = 5) # perform k-means clustering on a data matrix
pen3_kmeans
```

Result of 5 Clusters with 4 variables.
```{r}
tidy(pen3_kmeans)
```
Data visualzation
```{r}
penguins %>%
  select(species:body_mass_g) %>%
  drop_na() -> tmp
augment(pen3_kmeans, tmp) %>%
  ggplot(aes(bill_length_mm, bill_depth_mm, color = body_mass_g, size = flipper_length_mm)) +
  geom_point() +
  theme_bw() +
  facet_wrap(~.cluster)
```

# Elbow Chart
g. Construct an Elbow Chart to find an appropriate number of clusters for the data set
- We need to visualize the total within-cluster sum of squares with each k
- According to the Elbow Chart, we can say the roughly appropriate number of clusters is k = 2.
```{r}
tibble(k = 1:10) %>%
  mutate(models = map(k, ~kmeans(pen3, centers = .x))) %>% #.x = k %>%
  mutate(tot.withinss = map_dbl(models, "tot.withinss")) -> many_kmenas_res
```

Plotting
```{r}
many_kmenas_res %>%
  ggplot(aes(k, tot.withinss)) +
  geom_point() +
  geom_line() + 
  theme_bw() +
  ggtitle("Total within-cluster sum of squares vs K") +
  geom_vline(xintercept = 2, color = "red") +
  geom_text(x = 4, y = 800, label = "Elbow Point, K = 2", 
            size = 6, colour = "red")
```

# References
- https://en.wikipedia.org/wiki/K-means_clustering
- https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/kmeans
- https://stackoverflow.com/questions/20256028/understanding-scale-in-r