---
title: "PCA in the tidyverse Framework"
author: "Yunting Chiu"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: cerulean
---
```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
We will in this lab explore how principal components are calculated and analyzed.\
We will be using the `concrete` data set from the {modeldata} package which is loaded with {tidymodels}.

# Compressive Strength of Concrete Mixtures Data
```{r, message=FALSE}
library(ggbiplot)
library(tidymodels)
data(concrete)
head(concrete)
dim(concrete)
```

We must delete all non-numeric columns from the data because the `prcomp()` function can only deal with numeric columns. But the variables of`concrete` are all numeric variables so we don't do that.
```{r}
length(map_dbl(concrete, mean)) == length(concrete)
```

Take a look on variable's standard deviation.
```{r}
map_dbl(concrete, sd)
```
# Principal Component Analysis

a. Calculate the PCA of the data
- `scale.: TRUE/FALSE`: tells R if it should scale the data. That is, divide the standard deviation of each column from the corresponding column. The default is `FALSE`.
```{r}
# we scale the variables to have standard deviation one
concrete_pca <- prcomp(~ . -compressive_strength, data = concrete, scale. = TRUE) 
concrete_pca

# NO scale the variables
concrete_pca_noscale <- prcomp(~ . -compressive_strength, 
                               data = concrete, scale. = FALSE,
                               rank. = 8) # rank is redundant
concrete_pca_noscale
```

There are quantities in `concrete_pca`.
```{r}
names(concrete_pca)
```

The `center` and `scale` components correspond to the means and standard deviations of the variables that were used for scaling prior to implementing PCA.
```{r}
concrete_pca$center
concrete_pca$scale
```

b. Explore the loading, eigenvalues, and final projection using the broom package.

`tidy(matrix = "u")` returns from the original space into principle components space.
```{r}
concrete_pca %>%
  tidy(matrix = "u") %>% 
  head()
```

`tidy(matrix = "loadings")` returns from the principle components space back into the original space. That is, return the rotation matrix.
```{r}
concrete_pca %>%
  tidy(matrix = "loadings") # v", "rotation", "loadings" or "variables":
```

`tidy(matrix = "pcs")` returns the information about the eigenvalues. Or we can say this way is how we look at the variance explained by each PC
```{r}
concrete_pca %>%
  tidy(matrix = "pcs") # "d", "eigenvalues" or "pcs": returns information about the eigenvalues.
```
# Data Visualization

c. Visualize the projections. Look at how the scaling of the variables changes the projection
```{r}
# scaled and not scaled
augment(concrete_pca, newdata = concrete) %>% 
  head()
augment(concrete_pca_noscale, newdata = concrete) %>% 
  head()
```

```{r}
# another way to solely generate outcomes
# predict(concrete_pca_noscale, newdata = concrete)
```

We can see the location of rotated observations in the first principal direction (PC1) space\
(Note: The PC2 axis is the second most important direction and it is orthogonal to the PC1 axis).
```{r}
concrete_pca %>%
  tidy(matrix = "loadings") %>%
  filter(PC ==1) %>%
  ggplot(aes(value, column)) +
  geom_col() +
  theme_bw()
```

To see the outcomes on each PCA
```{r}
concrete_pca %>%
  tidy(matrix = "loadings") %>%
  ggplot(aes(value, column)) +
  geom_col() +
  facet_wrap(~PC)
```

Only PC1 and PC2 are compared
```{r}
concrete_pca %>%
  tidy(matrix = "loadings") %>%
  filter(PC <= 2) %>%
  ggplot(aes(value, column)) +
  geom_col() +
  facet_wrap(~PC)
```

See the outcomes on PC1 and PC2 with no scaling
```{r}
concrete_pca_noscale %>%
  tidy(matrix = "loadings") %>%
  filter(PC <= 2) %>%
  ggplot(aes(value, column)) +
  geom_col() +
  facet_wrap(~PC)
```

The first component explains most of the variation in the data, which shows the largest variation.
```{r}
concrete_pca %>%
  tidy(matrix = "pcs") %>%
  ggplot(aes(PC, percent)) +
  geom_col() +
  theme_bw() +
  scale_x_continuous(breaks = 1:8) +
  scale_y_continuous(
    labels = scales::percent_format(),
    expand = expansion(mult = c(0, 0.01))
  )
```

```{r}
map_dbl(concrete, ~length(unique(.x)))
```

Look at the data in PC1 and PC2's coordinates, we can see `water`, `fine_aggregate`, and `fly_ash` are all contribute to PC1. 
```{r, warning=FALSE}
ggbiplot(concrete_pca)
```

d. Use the {recipes} package to calculate the principal components.
```{r}
set.seed(1234)
concrete_split <- initial_split(concrete, strata = "compressive_strength")
concrete_train <- training(concrete_split)
concrete_test <- testing(concrete_split)
```

- `threshold = 0.75` indicates that `step_pca` should generate enough components to capture 75 percent of the variability in the variables
- `step_pca` creates a specification of a recipe step that will convert numeric data into one or more principal components.
```{r}
rec_spec <- recipe(compressive_strength ~., data = concrete_train) %>%
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors(), threshold = 0.75)
```

```{r}
lm_spec <- linear_reg() %>%
  set_mode("regression") %>%
  set_engine("lm")
```

```{r}
pcr_wf <- workflow() %>%
  add_recipe(rec_spec) %>%
  add_model(lm_spec)
```

```{r}
pcr_fit <- fit(pcr_wf, data = concrete_train)
pcr_fit
```


```{r}
augment(pcr_fit, new_data = concrete_test) %>%
  rmse(truth = compressive_strength, estimate = .pred) # root mean squared error
```

```{r}
augment(pcr_fit, new_data = concrete_test) %>%
  ggplot(aes(compressive_strength, .pred)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "green") +
  coord_fixed() +
  theme_bw()
```

# References
- https://broom.tidymodels.org/reference/tidy.prcomp.html
- https://clauswilke.com/blog/2020/09/07/pca-tidyverse-style/
- https://www.datacamp.com/community/tutorials/pca-analysis-r
- https://recipes.tidymodels.org/reference/step_pca.html
