---
title: "Labs02 - Week 1"
author: "Yunting Chiu"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 1 - Simple linear regression
a. Load the data biomass and plot HHV as a function of carbon.
```{r}
library(tidymodels)
# or run console table with "remotes::install_github("tidymodels/parsnip")"
data("biomass") # in model.data package
head(biomass)

# plot
biomass %>%
ggplot(aes(x = carbon, y = HHV)) +
  geom_point()
```

b. Fit a simple linear regression model to the data. With HHV as the response and carbon as the predictor. Is it a good fit?
- 
```{r}
# create a parsnip specification
linear_reg() %>%
  set_mode("regression") %>%
  set_engine("lm") -> lm_spec

# fit the model
lm_spec %>%
  fit(HHV ~ carbon, data = biomass)  -> lm_fit

# see the summary table
lm_fit %>%
  pluck("fit") %>%
  summary()
# another way to call summary table
# lm_fit$fit %>% summary()
```

c. Use the model to predict what the HHV of samples with carbon = 10, 20, ..., 80.
```{r}

```

d. Produce diagnostics plots. You can use plot() on the $fit object to produce some diagnostics.
```{r}

```

# Exercise 2 - Multiple linear regression
a. Fit a linear regression model to the data. With HHV as the response and carbon and hydrogen as the predictor. How is the fit compared to the simple linear model?
```{r}

```

b. Fit a linear regression model to the data. With HHV as the response and all the molecules as the predictor. How is the fit compared to the previous models?
```{r}

```

