---
title: "Labs 02 - Week 1 Wednesday"
author: "Yunting Chiu"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 1 - Simple linear regression
a. Load the data `biomass` and plot `HHV` as a function of `carbon`.
```{r, message=FALSE}
library(tidymodels)
# or run in console table with "remotes::install_github("tidymodels/parsnip")"
data("biomass") # in model.data package
head(biomass)

# plot
biomass %>%
ggplot(aes(x = carbon, y = HHV)) +
  geom_point() 
```

b. Fit a simple linear regression model to the data. With HHV as the response and carbon as the predictor. Is it a good fit?

- Both intercept and slope p-value are significant, meaning that we have evidence to reject the null hypothesis in favor of the $b_0$ and $b_1$ are not equal to zero. Meantime, the large R-squared indicates we have 85 % variation in `HHV` explained by `carbon`. Thus, the model is a good fit.
```{r}
# create a model specification
linear_reg() %>%
  set_mode("regression") %>%
  set_engine("lm") -> lm_spec

# fit the model
fit(lm_spec, HHV ~ carbon, data = biomass)  -> simple_lm_fit

# see the summary table
simple_lm_fit %>%
  pluck("fit") %>%
  summary()

# another way to call summary table
# simple_lm_fit$fit %>% summary()
```

```{r}
# using `tidy` to make a table as a tibble
simple_lm_fit %>% 
  tidy()
```

```{r}
# another way to extract the r-square
simple_lm_fit %>% 
  glance() %>%
  pull(r.squared)
```

c. Use the model to predict what the HHV of samples with carbon = 10, 20, ..., 80.
```{r}
# create a new HHV data frame
new_hhv <- tibble(carbon = seq(from = 10, to = 80, by = 10))
new_hhv

predict(simple_lm_fit, new_data = new_hhv) %>%
  rename(predictHHV = .pred)
```

d. Produce diagnostics plots. You can use `plot()` on the $fit object to produce some diagnostics.
```{r}
par(mfrow = c(2, 2))
simple_lm_fit %>%
  pluck("fit") %>%
  plot() 
```

# Exercise 2 - Multiple linear regression
a. Fit a linear regression model to the data. With `HHV` as the response and `carbon` and `hydrogen` as the predictor. How is the fit compared to the simple linear model?

- The p-values of $b_0, b_1, b_2$ are totally significant, and the r-squared is 0.85. We can say that this model and the previous simple linear model are similar. In the cause of we have 85 % variation can explain the models and the predictor(s) is/ are significant.
```{r}
fit(lm_spec, HHV ~ carbon + hydrogen, data = biomass)  -> multiple_lm_fit

summary(multiple_lm_fit$fit)
```

b. Fit a linear regression model to the data. With `HHV` as the response and all the molecules as the predictor. How is the fit compared to the previous models?

- The mean square error of three models are similar. However, We can see that `oxygen` and `nitrogen` are not significant in the full model, meaning that we don't have evidence that $b_3$ and $b_4$ are not equal to zero.  Because all of the predictors of the simple and multiple models are in the significant level, we can conclude that these two models are superior to the full model.
```{r}
fit(lm_spec, HHV ~ carbon + hydrogen + oxygen + nitrogen + sulfur, data = biomass)  -> full_lm_fit
summary(full_lm_fit$fit)
```

```{r}
# find out the three model's MSE, which are close to 2
anova(simple_lm_fit$fit) %>% tidy()
anova(multiple_lm_fit$fit) %>% tidy
anova(full_lm_fit$fit) %>% tidy()
```

